---
title: "Data Wrangling and Husbandry"
author: "Sai Teja Saggurthi"
date: "4/30/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rvest)
library(broom)
library(choroplethr)
library(choroplethrMaps)
library(data.table)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>


---

## Dataset

Dataset --> Olympics data.

* Data in CSV files. 
(https://www.kaggle.com/rio2016/olympic-games/data)

* Scraped data from Wikipedia. 
(https://en.wikipedia.org/wiki/
2016_Summer_Olympics_medal_table)

* Scraped data from Olympic database. 
(http://www.theolympicdatabase.nl/home)

Two data sources contain 2016 Olympics data and the other one has overall medals obtained by 
each country from start of the Olympics till 2016.

--- 

## Scraping and Loading Data

```{r warning=FALSE, echo=FALSE, message=FALSE}
athletes <- read_csv("E:/R_files_data_wrangling/olympic-games/athletes.csv")
athletes <- athletes %>% mutate(total = gold + silver + bronze)
countries <- read_csv("E:/R_files_data_wrangling/olympic-games/countries.csv")
colnames(countries)[2] <- "nationality"
```


```{r}
athletes %>% head(5)
countries %>% head(5)
```

---

Scraped the 2016 Olympics medal table from wikipedia and cleaned the data.

```{r warning=FALSE, message=FALSE}
url_wiki <- "https://en.wikipedia.org/wiki/2016_Summer_Olympics_medal_table"
medal_table2016 <- url_wiki %>% read_html() %>% 
  html_nodes("table") %>% .[2] %>% html_table(fill = TRUE) %>% data.frame()
medal_table2016 <- slice(medal_table2016, 1:86)
colnames(medal_table2016)[2] <- "Country"

```

```{r}
medal_table2016$Country <- str_replace_all(medal_table2016$Country, "\\([A-Z]{3}\\)", "")
medal_table2016$Country <- str_replace(medal_table2016$Country, "United States", "United States of America")
medal_table2016$Country <- str_replace(medal_table2016$Country, "Great Britain", "United Kingdom")
```

2016 Medal table

```{r}
medal_table2016 %>% head()
```

---

Scraping...

Overall Medal Table

```{r warning=FALSE, message=FALSE}
url_olympic <- "http://www.theolympicdatabase.nl/olympic/medals/countries"
medal_overall <- url_olympic %>% read_html() %>% 
  html_nodes("table") %>% .[[5]] %>% html_table(fill = TRUE) %>% data.frame()
medal_overall <- medal_overall[,c('X1','X3','X4','X6','X8')]
medal_overall <- medal_overall[-1,]
colnames(medal_overall) <- c("Rank","Country","Gold","Silver","Bronze")
medal_overall <- medal_overall %>% 
  mutate(Total = as.numeric(Gold) + as.numeric(Silver) + as.numeric(Bronze))
medal_overall <- slice(medal_overall, 1:135)
medal_overall %>% head()
```

---

```{r warning=FALSE, message=FALSE, echo=FALSE}
medal_overall$Country <- str_replace(medal_overall$Country, "United States", "United States of America")
medal_overall$Country <- str_replace(medal_overall$Country, "Great Britain", "United Kingdom")
medal_overall$Country <- str_replace(medal_overall$Country, "Soviet Union", "Russia")
```

---

Male to Female Ratio in 2016 Olympics 

```{r}
ggplot(athletes, aes(sex, fill=sex)) + geom_bar() +
  ggtitle("Bar Plot displaying number of male and female athletes") + xlab("Sex") +
  ylab("Count")
```

---

Total number of athletes for each sport

```{r}
ggplot(athletes, aes(sport)) + geom_bar() + 
  ggtitle("Bar Plot depicting various sports") + xlab("Sports") + ylab("Count") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1))
```

---

For each sport in the olympics which country is producing highest 
number of athletes.

```{r}
at <- athletes %>% group_by(sport, nationality) %>% summarise(total=n())
f<-at %>% filter(total == max(total))
ggplot(f, aes(sport,total)) + geom_col(aes(fill = nationality)) + xlab("Sport") +
  ylab("Number of athletes") + 
  ggtitle("Plot showing highest number of athletes for each sport by country") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1))
```

---


```{r warning=FALSE, message=FALSE, echo=FALSE}
atN <- athletes %>% group_by(nationality) %>% summarise(total=n())
country_gdp_pop <- countries %>% inner_join(atN)
colnames(country_gdp_pop)[5] <- "ath"
country_ath <- select(country_gdp_pop, country, ath)
country_ath$country <- sapply(country_ath$country, tolower)
country_ath$country <- gsub("\\*$", "", country_ath$country)
colnames(country_ath) <- c("region", "value")
#country_choropleth(country_ath, "Distribution of athletes across the world", 
#                   legend = "Number of athletes")
```

---

Weight distribution of players for each sport

```{r warning=FALSE}
ggplot(athletes, aes(sport, weight)) + geom_boxplot() +
  ggtitle("Box Plot depicting sports vs weight") + xlab("Sports") + ylab("Weight") +
  coord_flip()
```

---

Mean height vs weight for all the sports.

Lowest heights and weights are from Gymnastics. 

And, highest values are recorded from Basketball.

```{r warning=FALSE}
atT <- athletes %>% group_by(sport) %>% summarise(total=n())
atH <- athletes %>% group_by(sport) %>% summarise(mh=mean(height, na.rm = TRUE))
atW <- athletes %>% group_by(sport) %>% summarise(mw=mean(weight, na.rm = TRUE))
athl <- cbind(atH, atW, atT)
ggplot(athl, aes(mh, mw, color=sport, size=total)) + geom_point() + xlab("Height") + 
  ylab("Weight") + ggtitle("Height Vs Weight for each Sport ") +
  scale_size_continuous(range = c(3, 7))
```

---

Top 10 countries with Gold medals

```{r echo=FALSE}
top10_gold <- medal_table2016[order(medal_table2016$Gold, decreasing = TRUE),] %>% head(10)
top10_silver <- medal_table2016[order(medal_table2016$Silver, decreasing = TRUE),] %>% head(10)
top10_bronze <- medal_table2016[order(medal_table2016$Bronze, decreasing = TRUE),] %>% head(10)
```


```{r}
ggplot(top10_gold, aes(Country, Gold, fill=Country)) + geom_col(aes(alpha=0.2)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Top 10 countries with gold medals")
```

---

Top 10 countries with Silver medals


```{r}
ggplot(top10_silver, aes(Country, Silver, fill=Country)) + geom_col(aes(alpha=0.2)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Top 10 countries with silver medals")
```

---

Top 10 countries with Bronze medals

```{r}
ggplot(top10_bronze, aes(Country, Bronze, fill=Country)) + geom_col(aes(alpha=0.2)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Top 10 countries with bronze medals")
```

---

AthleteRatio - Ratio of number of athletes to population.

Top 10 nations with highest AthleteRatio

```{r}
country_gdp_pop <- country_gdp_pop %>% mutate(athleteRatio = ath/population)

country_gdp_pop[order(country_gdp_pop$athleteRatio, decreasing = TRUE),] %>% head(10)
```

---

Top 10 nations with lowest AthleteRatio

```{r}
country_gdp_pop[order(country_gdp_pop$athleteRatio),] %>% head(10)
```

---

Does gdp_per_capita of a country have an effect on number of athletes produced by a country?

```{r warning=FALSE}
ggplot(country_gdp_pop, aes(gdp_per_capita, ath)) + geom_point() + geom_smooth() +
  xlab("GDP Per Capita") + ylab("Athletes") + ggtitle("gdp per capita vs number of athletes")
```

---

Medals distribution for 2016 Olympics

```{r warning=FALSE, message=FALSE}
country_medal <- select(medal_table2016, Country, Total)
country_medal$Country <- sapply(country_medal$Country, tolower)
country_medal$Country <- gsub("\\*$", "", country_medal$Country)
trim.trailing <- function (x) sub("\\s+$", "", x)
country_medal$Country <- trim.trailing(country_medal$Country)
colnames(country_medal) <- c("region", "value")
country_choropleth(country_medal, "Medal density across the world in 2016 olympics",
                   legend = "Number of medals")
```

---

Medals distribution for overall Olympics

```{r warning=FALSE, message=FALSE}
medal_overall_new <- read_csv("E:/R_files_data_wrangling/olympic-games/medalData.csv")
country_overallmed <- select(medal_overall_new, Country, Total)
country_overallmed$Country <- sapply(country_overallmed$Country, tolower)
colnames(country_overallmed) <- c("region", "value")
country_choropleth(country_overallmed, "Medal density across the world in olympics",
                   legend = "Number of medals")
```

---

## Thank You

